### 2013-01-05 talk.tex
version of talk from Barry.

\documentclass[12pt]{beamer}

\setbeamertemplate{navigation symbols}{\insertbackfindforwardnavigationsymbol} 
\mode<presentation>
\usetheme{Madrid}
\usecolortheme{wolverine}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}

\usepackage{amssymb}
\usepackage[cmtip,all]{xy}

\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{url}
\usepackage[all]{xy}

\newtheorem*{thm}{Theorem}
\newtheorem*{lem}{Lemma}
\newtheorem*{rem}{Remark}
\newtheorem*{cor}{Corollary}
\newtheorem*{cor1}{Corollary 1}
\newtheorem*{cor2}{Corollary 2}
\newtheorem*{conj}{Conjecture}
\newtheorem*{prop}{Proposition}
\newtheorem*{STKconj}{Conjecture $\ST(K)$}

\theoremstyle{definition}
\newtheorem*{defn}{Definition}

\newtheorem*{exa}{Example}
\newtheorem*{exs}{Examples}

\newfont{\cyrr}{wncyr10}
\def\Sh{\mbox{\cyrr Sh}}
\def\S{\mathcal S}
\def\Z{\mathbb{Z}}
\def\Q{\bf {Q}}
\def\F{\mathbb{F}}
\def\R{\mathbb{R}}
\def\P{\mathbb{P}}
\def\x{\mathbf{x}}
\def\sp{{\rm Spec}}

\def\Zp{\Z_p}

\def\Fp{\F_p}
\def\Ftwo{\F_2}
\def\K{\mathcal{K}}
\def\cH{\mathcal{H}}
\def\A{\mathcal{A}}
\def\E{\mathcal{E}}
\def\O{\mathcal{O}}
\def\cR{\mathcal{R}}
\def\cS{\mathcal{S}}
\def\cP{\mathcal{P}}
\def\cN{\mathcal{N}}
\def\X{\mathcal{X}}
\def\ld{\mathcal{h}}
\def\rd{\mathcal{i}}

\def\mf{\mathfrak{f}}
\def\l{\mathfrak{l}}
\def\m{\mathfrak{m}}
\def\p{\mathfrak{p}}
\def\q{\mathfrak{q}}
\def\a{\mathfrak{a}}
\def\D{\mathfrak{d}}
\def\Pp{\mathfrak{P}}

\def\k{\Bbbk}

\def\Hom{\text{Hom}}
\def\Gal{\text{Gal}}
\def\rk{\text{rank}\,}
\def\ord{\text{ord}}
\def\ab{\text{ab}}
\def\tors{\text{tors}}
\def\coker{\text{coker}}
\def\Aut{\text{Aut}}
\def\Sel{\text{Sel}}
\def\End{\text{End}}
\def\Res{\text{Res}}
\def\Frob{\text{Frob}}
%\def\f{\text{f}}
\def\wc{\text{WC}}
\def\GL{\text{GL}}
\def\sign{\text{sign}}
\def\ST{\Sh\text{T}_2}
\def\jac{\text{Jac}}
\def\new{\text{new}}

\def\N{\mathbf{N}}

\def\too{\longrightarrow}
\def\map#1{\;\xrightarrow{#1}\;}
\def\isom{\xrightarrow{\sim}}
\def\hookto{\hookrightarrow}
\def\onto{\twoheadrightarrow}
\def\dirsum#1{\underset{#1}{\textstyle\bigoplus}}
\def\tens#1{\underset{#1}{\textstyle\bigotimes}}
%\def\cf{\mathrm{cond}_{\mathrm{f}}}
\def\cf{\mathrm{cond}}
\def\ls{\Omega}
\def\aug#1{{#1}^0}
%\def\Hf{H^1_\f}
\def\dens{\mathrm{density}}

\def\bmu{\boldsymbol{\mu}}

\title[How explicit is the Explicit Formula?] 
{How explicit is the Explicit Formula?}

\author[Barry Mazur and William Stein]{ Barry Mazur and William Stein\vskip30pt  {\it Rough notes for our combined talk at the  AMS Special Session on Arithmetic Statistics}}

%\institute[UCI]{\includegraphics[height=.3in]{formal-288.pdf}}

%\date{}
%\date[July 6, 2012]

\setbeamertemplate{headline}{}
\setbeamertemplate{footline}
{
  \leavevmode%
  \hbox{%
  \begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,center]{author in head/foot}%
    \usebeamerfont{author in head/foot}\insertshortauthor%~~(\insertshortinstitute)
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,center]{title in head/foot}%
    \usebeamerfont{title in head/foot}\insertshorttitle
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,right]{date in head/foot}%
    \usebeamerfont{date in head/foot}\insertshortdate{}\hspace*{4em}
    \insertframenumber{} / \inserttotalframenumber\hspace*{2ex} 
  \end{beamercolorbox}}%
  \vskip0pt%
}

\begin{document}

\begin{frame}
\titlepage
\end{frame}
\begin{frame}\vskip30pt 
{\Large The`Explicit Formulas' in analytic number theory deal with  {\it arithmetically interesting quantities}, often given as partial sums---the summands  corresponding to primes $p$---up to some cutoff value, $X$. We'll call them {``{\it Sums\ of\ local\ data},"}
\vskip40pt
\centerline{Again:}}
\end{frame}
\begin{frame}\vskip20pt
{\Large 
   A  {``{\it Sum\ of\ local\ data},"} is a sum of contributions for each prime $ p \le X$:

  $$\delta(X):=\ \ \  \sum_{p\le X}G(p)$$

  where the rules of the game require the value $G(p)$ to be determined by only {\it local }considerations at the prime $p$.} \end{frame}\begin{frame}\vskip20pt
{\Large

We will be concentrating on  {\it sums\ of\ local\ data}  attached to elliptic curves without CM over ${\Q}$,   $$\delta_E(X):=\sum_{p\le X}g_E(p)$$ where the weighting function $$p \mapsto g_E(p)$$   is a specific function of $p$ and $a_E(p)$, the $p$-th Fourier coefficient of the eigenform of weight two parametrizing the elliptic curve.}  \end{frame}\begin{frame}\vskip20pt
{\Large

  We will  be interested in  {\it issues of bias.} \vskip40pt  \centerline{\bf Weighted Biases}} \end{frame}\begin{frame}\vskip20pt
{\Large  \vskip20pt

Our Aim: to examine computations of these biases, following the classical `Explicit Formula," and the work of:

\vskip40pt
Sarnak, \ Granville,\  Rubenstein,\ Watkins,\vskip30pt \hskip40pt Martin,\ Fiorilli,\ Conrey-Snaith$\dots$ } \end{frame}\begin{frame}\vskip20pt
{\Large  \vskip20pt 
 For our elliptic curves $E$,\vskip20pt
 {\it ROUGHLY}---half the Fourier coefficients  $a_E(p)$ are positive and half negative.  } \end{frame}\begin{frame}\vskip20pt
{\Large \vskip20pt  That is: there are roughly as many $p$'s for which the number of rational points of $E$ over ${\bf F}_p$ is \vskip20pt \centerline{\it greater than $p+1$}  \vskip30pt \centerline{as there are primes for which it is} \vskip20pt \ \centerline{\it less than $p+1$}.       }  \end{frame}

\begin{frame}
 \vskip40pt {\Large   So let's study  finer statistical issues related to this symmetric distribution.   For example,  we can ask the {\it raw question:}  which of these classes of primes are winning the race, and how often? I.e., what can one say about   }  \end{frame}

\begin{frame}
 \vskip40pt {\Large  
 $$\Delta_E(X) =$$  \vskip20pt $$ \#\{p \ {\rm such\ that\ } |E({\bf F}_p| > p+1\}$$ \vskip10pt \centerline{minus} \vskip10pt  $$ \#\{p \ {\rm such\ that\ } |E({\bf F}_p| < p+1\}?$$}  \end{frame}

\begin{frame}
 \vskip40pt {\Large  Equivalently, putting:\begin{itemize} \item $\gamma_E(p)=0$ if $p$ is a bad or supersingular prime for $E$ and\vskip20pt \item i$\gamma_E(p)= -1$ if $E$ has more than $p+1$  ${\bf F}_p$-rational points, and \vskip20pt \item $\gamma_E(p) = +1$ if less.\end{itemize}}\end{frame}\begin{frame}\vskip20pt
{\Large \vskip40pt 
 So:\vskip20pt
   $$\Delta_E(X): =\sum_{p\le X}\gamma_E(p).$$

 More generally we might consider  weighting functions $p \mapsto g_E(p)$ that have the property that }  \end{frame}\begin{frame}\vskip20pt
{\Large \vskip40pt 
\begin{itemize} \item for all primes $p$, $g_E(p)$ is an {\it odd} function of the value  $a_E(p)$, and \vskip20pt \item the {\it sum\ of\ local\ data}  $$\delta_E(X):=\sum_{p\le X}g_E(p)$$ has---or can be convincingly conjectured to have---a finite  {\it mean}.\end{itemize}}\end{frame}
%\end{document}
\begin{frame}\vskip20pt
{\Large \vskip40pt 
Any such  $$p \mapsto  g_E(p)$$ represents a version of a `bias race'.
\vskip20pt
To illustrate specific features of the `Explicit Formula' we focus on three examples of such races for an elliptic curve $E$. }\end{frame}
  %\end{document}
\begin{frame}\vskip20pt
{\Large  \vskip20pt  Form  these 'sums of local data' ---\vskip20pt

We'll call them \vskip10pt  \centerline{the {\it raw},}\vskip10pt  \centerline{the {\it medium-rare}, and }\vskip10pt \centerline{the {\it well-done}}\vskip20pt 'sums of local data' }\end{frame}
  %\end{document}
\begin{frame}\vskip20pt
{\Large  \vskip20pt
 \centerline{\bf RAW:} \vskip20pt  $$\Delta_E(X): =\sum_{p\le X}\gamma_E(p)$$} \end{frame}
%\end{document}
\begin{frame}\vskip20pt
{\Large  \vskip20pt
 \centerline{\bf MEDIUM-RARE:}  \vskip20pt$${\mathcal D}_E(X):= {\frac{\log\ X}{\sqrt X}}\sum_{p \le X}{\frac{a_E(p)}{\sqrt p}}$$} \end{frame}
%\end{document}
\begin{frame}\vskip20pt
{\Large  \vskip20pt
  \centerline{\bf WELL-DONE:} \vskip20pt$${D}_E(X):= {\frac{1}{\log\ X}}\sum_{p \le X}{\frac{a_E(p)\log p}{ p}}$$} \end{frame}

 \begin{frame}\vskip20pt
{\Large  \vskip20pt

   The fun here is that there are clean conjectures for\vskip20pt the values of the {\it means} (relative to $dX/X$)\vskip20pt \centerline{---i.e., the {\it biases}---}\vskip20pt of the three `sums of local data' \vskip20pt and clean expectations of their {\it variances}:} \end{frame}

 \begin{frame}\vskip20pt
{\Large  \vskip20pt

    \begin{itemize}
   \item {\bf The well-done data:} the  mean is (conjecturally) $-r:=$  where $r= r_E$ is the {\it analytic rank} of $E$.  \vskip20pt

    \item {\bf The medium-rare data:} the  mean is  (conjecturally)  $1-2r$ and   \end{itemize} } \end{frame}

 \begin{frame}\vskip20pt
{\Large  \vskip20pt

      \centerline{\bf The raw data:}\vskip5pt  The  mean is  (conjecturally) $$
{\frac{2}{\pi}}- {\frac{16}{3\pi}}r$$ \vskip5pt \centerline{ + }  \vskip5pt $$ {\frac{4}{\pi}} \sum_{k=1}^{\infty}  (-1)^{k+1}\big[{\frac{1}{2k+1}} + {\frac{1}{2k+3}}\big]r({2k+1}).
$$ \centerline{  where }  } \end{frame}

 \begin{frame}\vskip20pt
{\Large  \vskip20pt $$r(n):= \ r_{f_E}(n)\ =$$ \\
$$=\ \  {\rm the\ order\ of\ vanishing\ of\ }$$

$$L(symm^nf_E, s)\ {\rm at}\ s=1/2,$$ \vskip20pt with $f_E:=$ the newform  corresponding to  $E$; and where  $s=1/2$ is the `central point.'  } \end{frame}

 \begin{frame}
{\Large  \vskip20pt  \centerline{ \bf Comments}
\vskip20pt {\bf(1)}  The (conjectured) distinction in the variances of the three formats.\vskip20pt
  \begin{itemize} \item   The raw data has {\it infinite variance} \item The medium-rare and well-done data have {\it finite variance}\end{itemize}} \end{frame}
 \begin{frame}
{\Large  \vskip20pt   {\bf(2)} The numbers \vskip20pt $$n \mapsto r_E(n)$$ \vskip20pt (for $n$ odd) conjecturally determine {\it all biases}!} \vskip20pt \centerline{\it Discuss}\end{frame}
 \begin{frame}\vskip20pt  
{\Large  \vskip20pt  \centerline{\it For example$\dots$}  \vskip20pt   If    $g(t)$ is a continuous function on $[-1,+1]$  with---appropriately defined---Fourier coefficients $\{c_n\}_n$, then the {\it mean} of the sum of local data,  $$G(X):= \sum_{p\le X}g\big(a(p)/2{\sqrt p}\big)$$  is conjecturally}\end{frame}
 \begin{frame}\vskip20pt  
{\Large  \vskip20pt    $$\sum_{n=1}^{\infty}  c_n\big(2r_E(n)+(-1)^n\big).$$\vskip20pt  
$$\{ {\rm{\it  \  Means\ of\ }} G(X){\rm{\it 's}}\} \ \ \  {\leftrightarrow}\ \ \  \{r_E(n){\rm{\it 's}}\}$$}\end{frame}
 \begin{frame}\vskip20pt
{\Large  \vskip20pt
  {\bf(3)} We have the beginnings of some data for those numbers, $n \mapsto r_E(n)$
 \vskip20pt
 but {\it nothing systematic}.  \vskip20pt
  {\bf(4)} And no firm conjectures yet.}\end{frame}

 \begin{frame}\vskip20pt
{\Large  \vskip20pt {\bf A very qualitative look at {\it the} Explicit Formula for our three `sums of local data'}

  $${\rm{\it Sum\ of\ local\ data\  }}  \ = $$
   \vskip20pt 
  $$  {\rm the\ {\it ``bias"}}\ + \  {\rm{\it  Oscillatory\ term}}\ + \  {\rm{\it  Error\ term}}.$$ }\end{frame}

 \begin{frame}\vskip20pt
{\Large  \vskip20pt   E.G., for the Well-done data,

$${D}_E(X):= {\frac{1}{\log\ X}}\sum_{p \le X}{\frac{a_E(p)\log p}{ p}}$$
the Explicit Formula gives ${D}_E(X)$ as  a sum of three contributions:

$$-r_E\ \ \  +\ \ \  S_E(X)\ \ \  +\ \ \  O(1/\log X)$$}\end{frame}

 \begin{frame}\vskip20pt
{\Large  \vskip20pt 
where the `{\rm{\it  Oscillatory\ term}}'  $S_E(X) $ is the wild card (even assuming GRH) and we take it to be the limit  ($Y \to {\infty}$) of these generalized trigonometric sums:  \vskip20pt 
 $$S_E(X,Y) = {\frac{1}{\log X}}\sum_{|\gamma| \le Y}{\frac{X^{i\gamma}}{i\gamma}},$$
}\end{frame}

 \begin{frame}\vskip20pt
{\Large  \vskip20pt  the sum being over the imaginary parts of the complex zeroes of  $L(f_E, s)\ {\rm at}\ s=1/2.$}\end{frame}

 \begin{frame}\vskip20pt
{\Large  \vskip20pt  
 It has been tentatively conjectured   that:

  $$\lim_{X,Y \to \infty}S_E(X,Y) = 0,$$  but for computations it would be good to know something more {\it explicit}.}\end{frame}

 \begin{frame}\vskip20pt
{\Large  \vskip20pt   
{\bf In sum, two issues needing conjectures, and computations:}
 \vskip20pt

 What should be conjectured about:

 \begin{enumerate}\item  the distribution of the $r_E(n)$'s?

 \vskip20pt \item the convergence of $\lim_{X,Y \to \infty}S_E(X,Y)$?\end{enumerate}}\end{frame}

\end{document}